Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation

### 2014-03-03 writedown/npt-neq/reply/reply.tex
Luigi's revise, send to Giovanni and Christof

\section*{The scientific issues from referee 1}

@@

The non-equilibrium responses in the paper are discussed in two
different senses: 1, the perturbation to the system is
time-independent.
The relaxation of the system to the new equilibrium
defined by the constant perturbation is non-equilibrium.
% However,
% after the new equilibrium is established, there is no non-equilibrium
% process in the system.
The first example of the manuscript is in this sense.  2, the
perturbation is time-dependent.  The second example in the manuscript
is in this sense.  The first sense is only a special case of the
sencond sense.  In the first sense, the system can be studied by
equilibrium methods such as the Markov state model (MSM). However, the
equilibrium methods are NOT applicable in the second sense.  The
present paper develops a non-equilibrium molecular dynamics simulation
method that is more general in the sense that it is applicable to the
aformentioned two senses. Therefore, the proposed method cannot be replaced
by the equilibrium methods such as MSM.

Another commenly used equilibrium method that is able to study the non-equilibrium
responses is the linear responses thoery (Green-Kubo relation).
We also use it to test the
conformational shift of alanine dipeptide under the electric field
(EF). The response function, take the time-dependent probability of
left-handed $\alpha$-helix (Conformation $C$ in the old manuscript. The notation has been changed to $\alpha_L$ in the updated manuscript.) for example, is defined as
\begin{align}\label{eqn:tmp1}
  P_C(t) = \langle\chi_{_C}(t)\rangle = \langle \chi_{_C} \rangle_0 -
  \beta \int_0^t ds\; M_e(t - s)\langle j(0)\cdot \chi_{_C}(t) \rangle_0
\end{align}
where $\chi_{_C}$ is the characteristic function of set $C$ that takes
the value of 1 for $(\phi,\psi)\in C$, and takes value 0
otherwise. $M_e(t)$ is the magnitude of the EF. In the case of
constant EF, $M_e(t) = t/t_{warm}$ for $0\leq t<t_{warm}$, and
$M_e(t) = 1$ for $t\geq t_{warm}$.
$j$ is the dissipative flux that is defined by
\begin{align}
  j = - \sum_{i=1}^N E_\infty q_i v_{i,x},
\end{align}
where $q_i$ is the partial charge of the $i$th atom, and $v_{i,x}$ is
the $x$ component of the velocity of the $i$th atom.  The notation
$\langle\cdot\rangle_0$ denotes the equilibrium ensemble average.
Eq.~\eqref{eqn:tmp1} can be found in the text books of statistical
mechanics (e.g.~M.~Tuckerman \emph{Statistical Mechanics: Theory and
  Molecular Simulation}, pp.~499).  We plot the correlation function
$\langle j(0)\cdot \chi_{_C}(t) \rangle_0$ against corrlation time $t$
in Fig.~\ref{fig:tmp1} of this reply. The value of the function is estimated from
two independent equilibrium simulation, and the trajactory of each is
of 1~$\mu$s. Please note that the total length of the equilibrium
trajactories is the same as the non-equilibrium simulation (2000
trajactories, each 1000~ps long).
% With the knowledge of
% non-equilibrium simulation, the correlation function is expected to
% vanish at roughly 500~ps, however, it does not vanish even at
% 1000~ps. This is not surprising, because
We find that the statistical error is even larger than the value of
corrlation function.  Therefore, the
overwhelming statistical uncertainty does not allow us to use it 
in Eq.~\eqref{eqn:tmp1} to calculate the
non-equilibrium averages. 
% , and we actually do not know when the
% corrlation function vanish .
Since the total computational effort of
the equilibrium simulation is the same as the non-equilibrium
simulation, we conclude that non-equilibrium simulatoin is more
precise than the conventional equilibrium approach (linear response) at the same computational cost.
A similar
observation has already been reported by Ciccotti and Jacucci, PRL
1975.
\\
% Another limit of the linear response theory is that it can only be
% used to calculate the non-equilibrium properties when the
% non-equilibrium purturbation to the system is small, and it is also
% impossible to tell \emph{a priori} if the purturbation is small enough
% that the response theory is valid. This issue also does not exist in
% our non-equilibrium simulation method.

\begin{figure}
  \centering
  \includegraphics{figs/fig-corr-meta.eps}
  \caption{The corrlation function $\langle j(0)\cdot \chi_{_C}(t)
    \rangle_0$ against time $t$. The error bars in the figure denote
    the statistical uncertainty at 95\% confidence level. }
  \label{fig:tmp1}
\end{figure}

@@

We have done an equilibrium simulation of 1~$\mu$s.  On the
trajactory, molecular conformations are recored every 500~ps,
therefore, in total 2000 equilibrium conformations are recoreded.  The
equilibrium probabilities of the conformations $A_1$, $A_2$, $B_1$,
$B_2$ and $C$ are 0.426 ($\pm 0.029$), 0.076 ($\pm 0.010$), 0.265
($\pm 0.013$), 0.182 ($\pm 0.019$) and 0.051 ($\pm 0.021$),
respectively.  We have run two addtional equilibrium simulations of
1~$\mu$s, and the probabilities of conformations are consistent with
the values listed above, which means that our new equilibrium
simulation is fully converged.  Comparing with our previous
equilibrium simulation (only 0.1~$\mu$s long),
% in which the
% probability of left-handed $\alpha$-helix is only 0.008,
the precision
of initial configuration sampling is substantially improved. All
simulations in the manuscript are reperformed according to the new set
of initial conformations, however, the strand of all time-dependent
observables are the same as the old simulation, and no change to the
conclusions is made.
% The difference between the new and the old
% results is no more than 0.05, which is basically within the
% statistical uncertainty (given in the new manuscript).

We also test the convergence of simulation result with respect to the
number of branching trajactories.  The time-dependent probabilities of
each conformation calculated from one branching trajactory is compared
with four braching trajactories (each of them is simulated with
different random seed for Langevin thermostat) in
Fig.~\ref{fig:tmp2} of this reply. The results are consistent, which means that the
sampling of the non-equilibrium trajactories is fully converged. In
summary, since the our non-equilibrium simulation includes the
sampling of both initial configurations and non-equilibrium
trajactories, the convergence of both aspects implies the convergence
of our simulation results.

The question on whether our non-equilibrium braching trajactories are
long enough so that the results are fully converged is a very subtle
question. There may exists an intrisic time-scale of 1 microsecond
long, so we cannot discover it by braching trajactories of only a few
nanoseconds.  What we can do by the proposed non-equilibrium simulation
method is to study time-dependent behaviours that are no longer than
the simulation time. We have modified the manuscript according to
make this point clear. A similar difficulty actualy exists for all equilibrium
simulations: how do we know whether an equilibrium simulation is fully converged,
especially when some rare events cannot be observed during the limited simualtion time.
\\
% We want to stress that the target of the non-equilibrium simulation is
% to test how the populations of conformations evolve under the external
% electric field

@@

We compared the simulation results of two different $R_{ex}$ values:
1.0~nm and 1.5~nm in Fig.~7 (pp.15) of the manuscript. The
$R_{ex}=1.0$~nm case was simulated in periodic cubic simulation box
of length $L=2.7$~nm and 4.0~nm.  The $R_{ex}=1.5$~nm case was
simulated in cubic box of length $L=4.0$~nm. Fig.~7 of the
manuscript demonstrates that all of the results are consistent. That
means $R_{ex}=1.0$~nm is already large enough, and a larger $R_{ex}$
would not change the results.\\

@@

The periods 10, 40~ps are roughly the same as the first two shortest
time scales of probability fluxes observed in the constant EF case,
see Table~I of the manuscript. We tried to simulation with period of
500~ps, which is the same as the longest timescales of the constant EF
case, however, the required simulation time is so long that exceed our
computational capability. Therefore, we made a compromise that only
the period of 200~ps is studied. The typical frequency of the microwave
irradiation is 2450~MHz, which corresponds to period of 400~ps. It is
the same order of magnitude as the $T_P=200$~ps studied by us.
\\

@@

Yes, the definition of the probability fluxes is similar to the
kinetic rate (we use joint probabilities in (6) rather than
conditional probabilities). However, we want to stress that, in genarl,
the probability fluxes calculated in the manuscript should not be
interpreted as a kinetic rate, and one cannot replace the
non-equilibrium molecular dynamics by a kinetic analysis. The main
reason is two folds:
Firstly, for a non-equilibrium system, the probability fluxes are
time-dependent, therefore, the lag-time $\Delta t$ in the definition (Eq.~(6))
should be small comparing with the typical time-scale of conformation change.
Secondly, for a small lag time $\Delta t$ (1~ps in the manuscript), the conformational
dynamics of biomolecules is generally not
Markovian (see e.g.~C.~Sch\"utte et.~al.~JCP 2011 and J.-H.~Prinz et.~al.~JCP 2011).
Since the basic assumption -- Markovianity -- of the kinetic analysis
may not be satisfied, a direct
application of the kinetic analysis in the non-equilibrium cases is questionable. Although one may
gain usefuly information from the kinetic analysis, the reliability
and the effectiveness of this approach should be carefully checked case
by case.  Since the main goal of our research is to propose a
non-equilibrium molecular dynamics simulation method, the numerical
results are not compared with a kinetic analysis.

A comment on this direction has been added to the end of section III.B.
\\

@@

Please see the reply to the seocond question concerning the convergence
of our simulation results. We believe that the better initial
state sampling and the new convergence check with respect to the
number of branching trajectories will strengthen the reliablity
of our method.
% manuscript and

@@

\section*{The scientific issues from referee 2}

@@

The frequency investigated in our manuscript is from 5~GHz to 100~GHz,
the largest of which is still one order of magnitude smaller than the
terahertz spectroscopy. Therefore, we find our simulation results
cannot be directly compared with the terahertz experiments.\\

@@

Yes, a comment and comparison on this direction has been added to the
manuscript.\\

@@

\emph{
It is suggested to adapt the labeling of the conformational substates
of alanine dipeptide to that used in the literature ($C_7^{eq}$, $\alpha_L$,
$C_7^{ax}$, etc.) - see e.g. Caflisch et al., JCP 1999.
}\\

We have changed the labelling of conformations.\\

@@

This is indeed a very interesting point. Although the Fig.11 is
plotted for the non-equilibrium oscillatory EF case, the
confromational transition pathways are qulitatively comparable to the
previous works on the transition pathways studied in equilibrium.
Both the connectivity of conformations and the relatively strength of
the transitions are similar. A short comment has been added to the
manuscript.

The transition fluxes can be used to calculate the free energy only in
the equilibrium case. In the non-equilibrium cased, especially the
oscillatory EF case, the conformational kinetics of the system is generally
not Markovian. Therefore, the standard kinetic analysis cannot be used
here.

>>>>

Dear Editor,\\

Below you find our reply to the the referees. We apologize for the long time needed for the reply due to the holiday of one of the authors.  The request of the referees implied a rather massive additional computational effort. The result of such effort is highly satisfactory and the validity of the paper is now much higher. We are now confident that the paper can be accepted for publication.\\

Best Regards\\

Han Wang (on behalf of all authors) 
\section*{Reply to referee 1}

@@

We understand the concern of the referee and in the revised version of our manuscript have performed a numerical test.

We have employed equilibrium simulations to obtain the non-equilibrium
responses within the framework of linear responses theory (Green-Kubo relation).
With this approach we studied the conformational shift of alanine dipeptide under the electric field
(EF). As response function we have considered the time-dependent probability of
left-handed $\alpha$-helix (Conformation $C$ in the old manuscript. The notation has been changed to $\alpha_L$ in the updated manuscript.) This is defined as
\begin{align}\label{eqn:tmp1}
  P_C(t) = \langle\chi_{_C}(t)\rangle = \langle \chi_{_C} \rangle_0 -
  \beta \int_0^t ds\; M_e(t - s)\langle j(0)\cdot \chi_{_C}(t) \rangle_0
\end{align}
where $\chi_{_C}$ is the characteristic function of set $C$ that takes
the value of 1 for $(\phi,\psi)\in C$, and takes value 0
otherwise. $M_e(t)$ is the magnitude of the EF. In the case of
constant EF, $M_e(t) = t/t_{warm}$ for $0\leq t<t_{warm}$, and
$M_e(t) = 1$ for $t\geq t_{warm}$.
$j$ is the dissipative flux that is defined by
\begin{align}
  j = - \sum_{i=1}^N E_\infty q_i v_{i,x},
\end{align}
where $q_i$ is the partial charge of the $i$th atom, and $v_{i,x}$ is
the $x$ component of the velocity of the $i$th atom.  The notation
$\langle\cdot\rangle_0$ denotes the equilibrium ensemble average.
 We then plot the correlation function
$\langle j(0)\cdot \chi_{_C}(t) \rangle_0$ against correlation time $t$
in Fig.~\ref{fig:tmp1} of this reply. The value of the function is estimated from
two independent equilibrium simulations, and the trajectory of each is
of 1~$\mu$s. The total length of the equilibrium
trajectories is the same as the non-equilibrium simulation (2000
trajectories, each 1000~ps long).

We find that the statistical error is even larger than the value of
correlation function.  Therefore, the
overwhelming statistical uncertainty does not allow us to use the equilibrium approach 
in Eq.~\eqref{eqn:tmp1} to calculate the
non-equilibrium averages. 
% , and we actually do not know when the
% corrlation function vanish .
Since the total computational effort of
the equilibrium simulation is the same as the non-equilibrium
simulation, we conclude that non-equilibrium simulation is more
accurate and efficient than the conventional equilibrium approach  at the same computational cost.
It must be noticed that a similar
observation has already been reported by Ciccotti and Jacucci, (Phys.~Rev.~Lett.~35, 789, 1975). We report the description above in the appendix of the manuscript.
\\
\begin{figure}
  \centering
  \includegraphics{figs/fig-corr-meta.eps}
  \caption{The correlation function $\langle j(0)\cdot \chi_{_C}(t)
    \rangle_0$ against time $t$. The error bars in the figure denote
    the statistical uncertainty at 95\% confidence level. }
  \label{fig:tmp1}
\end{figure}

@@

We have done an equilibrium simulation of 1~$\mu$s.  Along the
trajectory, molecular conformations are recorded every 500~ps,
therefore, in total 2000 equilibrium conformations are extracted.  The
equilibrium probabilities of the conformations $A_1$, $A_2$, $B_1$,
$B_2$ and $C$ are 0.426 ($\pm 0.029$), 0.076 ($\pm 0.010$), 0.265
($\pm 0.013$), 0.182 ($\pm 0.019$) and 0.051 ($\pm 0.021$),
respectively.  We run two additional equilibrium simulations of
1~$\mu$s, and the probabilities of conformations are consistent with
the values listed above, which means that our ``new'' equilibrium
simulation is fully converged.  Comparing with our previous
equilibrium simulation (only 0.1~$\mu$s long),
% in which the
% probability of left-handed $\alpha$-helix is only 0.008,
the precision
of initial configuration sampling is substantially improved. All
simulations in the manuscript are re-performed according to the new set
of initial conformations, however nothing substantially changed compared to the original results.\\

We have also tested the convergence of simulation results with respect to the
number of branching trajectories.  The time-dependent probabilities of
each conformation calculated from one branching trajectory is compared
with four branching trajectories (each of them is simulated with
different random seeds for Langevin thermostat) in
Fig.~\ref{fig:tmp2} of this reply. The results are consistent with the original ones, which means that the
sampling of the non-equilibrium trajectories is fully converged. In
summary, since our non-equilibrium simulation includes the
sampling of both initial configurations and non-equilibrium
trajectories, the convergence of both  implies the convergence
of our simulation results. 
The updated results are added in the paper.\\

The question on whether our non-equilibrium branching trajectories are
long enough so that the results are fully converged is a very subtle
question. There may exists an intrinsic time-scale longer than our simulation time, however this is obvious and implicitly imply that the validity of our results is restricted to the time scale simulated. 
\\

@@

We compared the simulation results of two different $R_{ex}$ values:
1.0~nm and 1.5~nm in Fig.~7 (pp.15) of the manuscript. The
$R_{ex}=1.0$~nm case was simulated in periodic cubic simulation box
of length $L=2.7$~nm and 4.0~nm.  The $R_{ex}=1.5$~nm case was
simulated in cubic box of length $L=4.0$~nm. Fig.~7 of the
manuscript demonstrates that all the results are consistent. This
means $R_{ex}=1.0$~nm is already large enough, and a larger $R_{ex}$
would not change the results.\\

@@

The periods 10, 40~ps are roughly the same as the first two shortest
time scales of probability fluxes observed in the constant EF case,
see Table~I of the manuscript. We tried to simulation with period of
500~ps, which is the same as the longest timescales of the constant EF
case, however, the required simulation time is prohibitive at this stage. Therefore, we made a reasonable compromise and considered only
a period of 200~ps. The typical frequency of the microwave
irradiation is 2450~MHz, which corresponds to period of 400~ps. It is
the same order of magnitude as the $T_P=200$~ps studied by us.
We have added a comment along these lines in the revised manuscript.
\\

@@

Yes, the definition of the probability fluxes is similar to the
kinetic rate (we use joint probabilities in (6) rather than
conditional probabilities). However, we want to stress out that, in general,
the probability fluxes calculated in the manuscript should not be
interpreted as a kinetic rate, and one cannot replace the
non-equilibrium molecular dynamics by a kinetic analysis. The main
reasons are:
Firstly, for a non-equilibrium system, the probability fluxes are
time-dependent, therefore, the lag-time $\Delta t$ in the definition (Eq.~(6))
should be small compared with the typical time-scale of conformation change.\\
Secondly, for a small lag time $\Delta t$ (1~ps in the manuscript), the conformational
dynamics of biomolecules is generally not
Markovian (see e.g.~C.~Sch\"utte et.~al.~JCP 2011 and J.-H.~Prinz et.~al.~JCP 2011).
Since the basic assumption -- Markovianity -- of the kinetic analysis
may not be satisfied, a direct
application of the kinetic analysis in the non-equilibrium cases is questionable. Although one may
gain useful information from the kinetic analysis, the reliability
and the effectiveness of this approach should be carefully checked case
by case.  Since the main goal of our research is to propose a
non-equilibrium molecular dynamics simulation method, the numerical
results are not compared with a kinetic analysis.\\

A comment along this direction has been added to the end of section III.B.
\\

@@

The referee is invited to look at the reply to the second question concerning the convergence
of our simulation results. We believe that a better initial
state sampling and the new convergence check with respect to the
number of branching trajectories will strengthen the reliability
of our method.
% manuscript and

@@

\section*{Reply to referee 2}

@@

The frequencies investigated in our manuscript span from 5~GHz to 100~GHz,
the largest of which is still one order of magnitude smaller than those normally used in  terahertz spectroscopy. Therefore, we find our simulation results, at this stage, cannot be directly compared with the terahertz experiments.
{To make fully aware the readers about the current limitation of our study, with respect to experimental connections, we have added the comment above in the manuscript.}
\\

@@

Yes, a comment and comparison along this direction has been added to the
manuscript.\\

@@

\emph{
It is suggested to adapt the labeling of the conformational substates
of alanine dipeptide to that used in the literature ($C7_{eq}$, $\alpha_L$,
$C7_{ax}$, etc.) - see e.g. Caflisch et al., JCP 1999.
}\\

We have changed the labeling of conformations.\\

@@

This is indeed a very interesting point.
Although the network presented Fig.11 is similar to those
discovered in the mentioned equilibrium studies,
regarding the map of the steady conformations and connections among them,
the physical meaning is instead quite different in our study:
The system is under the nonequilibrium conditions, and
the fluxes are explicitly induced by a physically realistic control, i.e.~EF, although the averaged population of the steady conformations do not change. In contrast,
such probability fluxes do not exist, as a result of an explicit physical action, in the equilibrium case.
A comment along such direction has been added to the
manuscript.\\

It must be also added that the transition fluxes can be used to calculate the free energy only in
the equilibrium case. In the non-equilibrium case, especially for the
oscillatory EF case, the conformational kinetics of the system is generally
not Markovian. Therefore, the standard kinetic analysis cannot be used
here.
